```{r}
library(ggplot2)
library(dplyr)
carbon <- read.csv("carbon-monitor.csv")
file.exists("carbon-monitor.csv")
country <- carbon$country
date <- carbon$date
sector <- carbon$sector
value <- carbon$value
timestamp <- carbon$timestamp

unique(carbon$sector)

filter(carbon, country == "Austria", between(date, "06/08/21", "15/08/21"), sector == "Power")
filter(carbon, value == "0")
class(timestamp)
class(sector)
str

mean(carbon$value)
class(carbon$timestamp)

as.numeric(timestamp)


## Replacing 0's ####

# Austria Power: 06/08/2021 - 15/08/2021 replaced by 16/08/21 - 25/08/2021 Mean
AustriaPowerMean <- carbon |>
  filter(country == "Austria" & sector == "Power" & between(timestamp, 1629068400, 1629846000)) |>
  summarize(mean_value = mean(value)) |>
  pull(mean_value)

# Austria Industry: 06/08/2021 - 13/08/2021 replaced by 14/08/2021 - 21/08/2021 Mean
AustriaIndustryMean <- carbon |>
  filter(country == "Austria" & sector == "Industry" & between(timestamp, 1628895600, 1629500400)) |>
  summarize(mean_value = mean(value)) |>
  pull(mean_value)



## New dataset that removes 0's and EU27 & UK
carbonNew <- carbon |>
  filter(country != "EU27 & UK") |>
  mutate(value = ifelse(country == "Austria" & sector == "Power" & value == 0, AustriaPowerMean,
                        ifelse(country == "Austria" & sector == "Industry" & value == 0, AustriaIndustryMean, value)))

## Seperating into years
carbon2019 <- filter(carbonNew, timestamp <= "1577750400")

filter(carbonNew, timestamp >= "1577750400")


# Quality Control
testing <- filter(carbonNew, country == "Latvia",
                             sector == "Power", value == "0")

latvia_data <- carbon %>%
  filter(country == "Latvia") %>%
  select(sector, value)
ggplot(latvia_data, aes(x = sector, y = value)) +
  geom_jitter() +
  labs(x = "Sector", y = "Emissions", title = "Carbon Emissions from Different Sectors in Latvia")

```

```{r}
## Population Density ####
density <- read.csv("population-density.csv")
library("dplyr")
location <- density$Location

filter(newDensity, location == "World")

densityNew <- density |>
  na.omit() |>
  filter(density$Location != "World")

?if_any
tester <- density %>%
  filter(if_any(everything(), ~ !is.na(.x)))

filter(density, location == "Croatia")

newDensity <- density %>%
  select(location, everything())
```

```{r}
## Playing around with plots

ggplot(carbonNew, aes(x = country, y = value)) +
  geom_count() +
  theme_minimal()

ggplot(carbon2019, aes(x = country, y = value)) + 
  geom_count() +
  theme_minimal()
?heatmap

carbonNew |>
  filter(country %in% c("Austria", "Germany", "Luxembourg", "Belgium", "France")) |>
  ggplot(aes(x = country, y = sector, fill = value)) +
  geom_tile() +
  labs(x = "Country", y = "Variable", title = "Heatmap of Different Values for Selected Countries") +
  scale_fill_viridis_c()  # Adjust the color scale if needed
```

```{r}
## Timetables
library(lubridate)
parse_date_time
```

#### Notes

-   Which 0's should I not remove?

    -   Latvia Industry

    -   Luxembourg Domestic Aviation January 2019, a lot of 2022

-   Latvia data is insignificant

-   EU27 & UK data

-   Why is it only European countries

-   What units is values in

-   Good dataset for population density

What role does population density have on different facets of carbon emissions?

-   Look into density of neighboring countries

-   neighboring small countries emissions might be impacted by larger countries

    -   sharing resources

-   some sectors might not scale to population density

    -   take into account geographic region

    -   locate largest carbon emission place

        -   ex. large international aviation in germany

            -   does nearby international aviation decrease while ground transport increases?

            -   
